Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-11-06 evaluation.tex
Typos.

In this section, we present our experience in running \thetool on two
different applications: \emph{Courseware}: a course registration
system described by~\cite{gotsmanpopl16}, and the \emph{TPC-C}
benchmark.

@@

\begin{enumerate}
\item  $I_1$: An enrollment record should always refer to an existing student and an existing course.
\item  $I_2$: The capacity (\C{c\_capacity}) of a course should always be a
  non-negative quantity.
\end{enumerate}
\noindent Both $I_1$ and $I_2$ can be violated under weak isolation.
$I_1$ can be violated, for example, when \C{deregister\_student} runs
concurrently with \C{enroll}, both at RC isolation. While the former
transaction removes the student record after checking that no
enrollments for that student exists, the latter transaction
concurrently adds an enrollment record after checking the student
exists.  Both can succeed concurrently, resulting in an invalid state.
Invariant $I_2$ can be violated by two \C{enroll}s, both reading
\C{c\_capacity}=1, and both (atomically) decrementing it, resulting in
\C{c\_capacity}=-1.  We ran \thetool on the courseware application
(Fig.~\ref{fig:courseware_code}) after annotating transactions with
their respective guarantees, and asserting $I = I_1 \wedge I_2$ as the
correctness condition. The guarantees $G_e$ and $G_d$ for \C{enroll}
and \C{deregister\_student} transactions, respectively, are shown below:
\begin{smathpar}
  \begin{array}{lcl}
    G_e(\stg,\stg') & \Leftrightarrow & \stg_s'=\stg_s
      \conj \exists\C{cid}.\exists\C{sid}.\\
    & & \hspace*{0.3in} \stg_c' = \stg_c \bind \lambda c.~
        \itel{c.\C{c\_id}=\C{cid}\\
    & & \hspace*{1.15in}}
          {\existsl(c',~c'.\C{id}=c.\C{id} \wedge
              c'.\C{c\_name}=c.\C{c\_name} \wedge c'.\C{c\_capacity}\ge0,
              ~\{c'\})\\
    & & \hspace*{1.15in}}
          {\{c\}}\\
    & & \hspace*{0.15in}\conj \stg_e = \stg_e' \bind \lambda e.~ 
        \itel{e.\C{e\_c\_id}=\C{cid} \wedge e.\C{e\_s\_id}=\C{sid}}
          {\emptyset}{\{e\}}\\
%   & & \hspace*{1.15in}} 
%         {\emptyset} {\{e\}}\\
    G_d(\stg,\stg') & \Leftrightarrow & \stg_c' = \stg_c \conj
      \stg_e' = \stg_e \conj \exists \C{sid}.~
      \itel{\forall(e \in \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
%   & & \hspace*{0.3in} \itel{\forall(e \in
%           \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s \bind \lambda s.~
           \itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s}\\
  \end{array}
\end{smathpar}
For the sake of this presentation we split $\stg$ into three disjoint
sets of records, $\stg_s$, $\stg_c$, and $\stg_e$, standing for
\C{Student}, \C{Course}, and \C{Enrollment} tables, respectively.
Observing that the set language $\SL$ (Sec.~\ref{sec:inference}),
besides being useful for automatic verification, also facilitates
succinct expression of transaction semantics, we define $G_e$ and
$G_d$ in a combination of FOL and $\SL$. $G_e$ essentially says that
the \C{enroll} transaction leaves the \C{Student} table unchanged,
while it may update the \C{c\_capacity} field of a \C{Course} record
to a non-negative value (even when it doesn't update, it is the case
that $c'.\C{c\_capacity}\ge0$, because $c'=c$, and $c\in\stg_c$, and
we know that $I_2(\stg_c)$). $G_e$ also conveys that \C{enroll} might
insert a new \C{Enrollment} record by stating that $\stg_e$, the
\C{Enrollment} table in pre-state, contains all records $e$ from
$\stg_e'$, the table in post-state, except when $e.\C{e\_c\_id}$ and
$e.\C{e\_s\_id}$ match \C{cid} and \C{sid}, respectively. The
guarantee $G_d$ of \C{deregister\_student} asserts that the
transaction doesn't write to \C{Course} and \C{Enrollment} tables. The
transaction might however delete a \C{Student} record bearing an
\C{id}=\C{sid} (formally, $\stg_s' = \stg_s \bind \lambda s.~
\itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}$), for some \C{sid} for
which no corresponding \C{Enrollment} records are present in the
pre-state (in other words, $\forall(e \in
\stg_e).~e.\C{e\_s\_id}\neq\C{sid}$).

>>>>

In this section, we present our experience in running \thetool on two
different applications: \emph{Courseware}: a course registration
system described by~\cite{gotsmanpopl16}, and the TPC-C benchmark.

@@

\begin{enumerate}
\item  $I_1$: An enrollment record should always refer to an existing student and an existing course.
\item  $I_2$: The capacity (\C{c\_capacity}) of a course should always be a
  non-negative quantity.
\end{enumerate}
\noindent Both $I_1$ and $I_2$ can be violated under weak isolation.
$I_1$ can be violated, for example, when \C{deregister\_student} runs
concurrently with \C{enroll}, both at RC isolation. While the former
transaction removes the student record after checking that no
enrollments for that student exists, the latter transaction
concurrently adds an enrollment record after checking the student
exists.  Both can succeed concurrently, resulting in an invalid state.
Invariant $I_2$ can be violated by two \C{enroll}s, both reading
\C{c\_capacity}=1, and both (atomically) decrementing it, resulting in
\C{c\_capacity}=-1.  We ran \thetool on the Courseware application
(Fig.~\ref{fig:courseware_code}) after annotating transactions with
their respective guarantees, and asserting $I = I_1 \wedge I_2$ as the
correctness condition. The guarantees $G_e$ and $G_d$ for \C{enroll}
and \C{deregister\_student} transactions, respectively, are shown below:
\begin{smathpar}
  \begin{array}{lcl}
    G_e(\stg,\stg') & \Leftrightarrow & \stg_s'=\stg_s
      \conj \exists\C{cid}.\exists\C{sid}.\\
    & & \hspace*{0.3in} \stg_c' = \stg_c \bind \lambda c.~
        \itel{c.\C{c\_id}=\C{cid}\\
    & & \hspace*{1.15in}}
          {\existsl(c',~c'.\C{id}=c.\C{id} \wedge
              c'.\C{c\_name}=c.\C{c\_name} \wedge c'.\C{c\_capacity}\ge0,
              ~\{c'\})\\
    & & \hspace*{1.15in}}
          {\{c\}}\\
    & & \hspace*{0.15in}\conj \stg_e = \stg_e' \bind \lambda e.~ 
        \itel{e.\C{e\_c\_id}=\C{cid} \wedge e.\C{e\_s\_id}=\C{sid}}
          {\emptyset}{\{e\}}\\
%   & & \hspace*{1.15in}} 
%         {\emptyset} {\{e\}}\\
    G_d(\stg,\stg') & \Leftrightarrow & \stg_c' = \stg_c \conj
      \stg_e' = \stg_e \conj \exists \C{sid}.~
      \itel{\forall(e \in \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
%   & & \hspace*{0.3in} \itel{\forall(e \in
%           \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s \bind \lambda s.~
           \itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s}\\
  \end{array}
\end{smathpar}
For the sake of this presentation we split $\stg$ into three disjoint
sets of records, $\stg_s$, $\stg_c$, and $\stg_e$, standing for
\C{Student}, \C{Course}, and \C{Enrollment} tables, respectively.
Observing that the set language $\SL$ (Sec.~\ref{sec:inference}),
besides being useful for automatic verification, also facilitates
succinct expression of transaction semantics, we define $G_e$ and
$G_d$ in a combination of FOL and $\SL$. $G_e$ essentially says that
the \C{enroll} transaction leaves the \C{Student} table unchanged,
while it may update the \C{c\_capacity} field of a \C{Course} record
to a non-negative value (even when it doesn't update, it is the case
that $c'.\C{c\_capacity}\ge0$, because $c'=c$, and $c\in\stg_c$, and
we know that $I_2(\stg_c)$). $G_e$ also conveys that \C{enroll} might
insert a new \C{Enrollment} record by stating that $\stg_e$, the
\C{Enrollment} table in pre-state, contains all records $e$ from
$\stg_e'$, the table in post-state, except when $e.\C{e\_c\_id}$ and
$e.\C{e\_s\_id}$ match \C{cid} and \C{sid}, respectively. The
guarantee $G_d$ of \C{deregister\_student} asserts that the
transaction doesn't write to \C{Course} and \C{Enrollment} tables. The
transaction might however delete a \C{Student} record bearing an
\C{id}=\C{sid} (formally, $\stg_s' = \stg_s \bind \lambda s.~
\itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}$), for some \C{sid} for
which no corresponding \C{Enrollment} records are present in the
pre-state (in other words, $\forall(e \in
\stg_e).~e.\C{e\_s\_id}\neq\C{sid}$).

### 2017-11-06 motivation.tex
Typos.

\begin{figure}
\centering
\begin{ocaml}
let new_order (d_id, c_id, item_reqs) = atomically_do @@ fun () ->
  let dist = SQL.select1 District (fun d -> d.d_id = d_id) in
  let o_id = dist.d_next_o_id in
  SQL.update(* UPDATE *) District 
            (* SET *)(fun d -> {d with d_next_o_id =d_next_o_id + 1})
            (* WHERE *)(fun d -> d.d_id = d_id );
  SQL.insert(* INSERT INTO *) Order (* VALUES *){o_id=o_id;  
            o_d_id=d_id; o_c_id=c_id; o_ol_cnt=S.size item_reqs; };
  foreach item_reqs @@ fun item_req ->
    let stk = SQL.select1(* SELECT * FROM *) Stock 
              (* WHERE *)(fun s -> s.s_i_id = item_req.ol_i_id 
                                  && s.s_d_id = d_id)(* LIMIT 1 *); 
    let s_qty' = if stk.s_qty >= item_req.ol_qty + 10 
                then stk.s_qty - item_req.ol_qty 
                else stk.s_qty - item_req.ol_qty + 91 in
        SQL.update Stock (fun s -> {s with s_qty = s_qty'}) 
                         (fun s -> s.s_i_id = item_req.ol_i_id);
        SQL.insert Order_line {ol_o_id=o_id; ol_d_id=d_id; 
                               ol_i_id=item_req.ol_i_id; ol_qty=item_req.ol_qty}

>>>>

\begin{figure}
\centering
\begin{ocaml}
let new_order (d_id, c_id, item_reqs) = atomically_do @@ fun () ->
  let dist = SQL.select1 District (fun d -> d.d_id = d_id) in
  let o_id = dist.d_next_o_id in
  begin
    SQL.update(* UPDATE *) District 
              (* SET *)(fun d -> {d with d_next_o_id = d.d_next_o_id + 1})
              (* WHERE *)(fun d -> d.d_id = d_id );
    SQL.insert(* INSERT INTO *) Order (* VALUES *){o_id=o_id;  
              o_d_id=d_id; o_c_id=c_id; o_ol_cnt=S.size item_reqs; };
    foreach item_reqs @@ fun item_req ->
      let stk = SQL.select1(* SELECT * FROM *) Stock 
                (* WHERE *)(fun s -> s.s_i_id = item_req.ol_i_id &&
                                     s.s_d_id = d_id)(* LIMIT 1 *) in
      let s_qty' = if stk.s_qty >= item_req.ol_qty + 10 
                  then stk.s_qty - item_req.ol_qty 
                  else stk.s_qty - item_req.ol_qty + 91 in
      SQL.update Stock (fun s -> {s with s_qty = s_qty'}) 
                       (fun s -> s.s_i_id = item_req.ol_i_id);
      SQL.insert Order_line {ol_o_id=o_id; ol_d_id=d_id; 
                             ol_i_id=item_req.ol_i_id; ol_qty=item_req.ol_qty}
  end
